---
title: "Lab 5A"
author: "Valentina Fender"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

library(dplyr)

lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv") |> 
  rename(Species = Race)
```


```{r blank original}
(lotr_wide <- lotr |> 
  pivot_wider(FILL_THIS_IN = c(Film, Gender), 
              FILL_THIS_IN = Species, 
              FILL_THIS_IN = Words)) 
```
 
# 2. Widen the data so that we see the words spoken by each species, by putting species as its own column.

```{r}
(lotr_wide <- lotr |> 
  pivot_wider(names_from = c(Film, Gender), 
              id_cols = Species, 
              values_from = Words))
```

^The code above is absolutely beautiful and perfect in every way. I love it so much.

`

```{r my attempts}
(lotr_w2 <- lotr |>
   pivot_wider(names_from = c(Film, Gender)))
```

```{r my attempts}
(lotr_w3 <- lotr |> 
  pivot_wider(names_from = c(Film, Gender), 
              id_cols = Species, 
              values_from = Words))
```



##Notes: pivot_wider() is the opposite of pivot_longer(): it makes a dataset wider by increasing the number of columns and decreasing the number of rows. It’s relatively rare to need pivot_wider() to make tidy data, but it’s often useful for creating summary tables for presentation, or data in a format needed by other tools.





































